---
title: Data analysis report - Root Insurance

# Use letters for affiliations
author:
  - name: Issac Lee
    affiliation: a
address:
  - code: a
    address: Department of Statistics & Actuarial Science, 241 Schaeffer Hall, Iowa City, Iowa 52242-1409

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
#date_subtitle: Published in *Journal of Statistical Software*, 2018

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Author and Author

# Place eg a DOI URL or CRAN Package URL here
doi_footer: "https://cran.r-project.org/package=YourPackage"

# Abstract
abstract: |
  This report illustrates how I approach the project and the decision making the process for Root insurance's data analysis project. The goal of this project is to find the best-matched OBDII trip data, which corresponds to a trip data from a user's smartphone. The data analysis process appears in the first chapter, and the instructional manual follows in the later chapter. The `R` code implementations follow Google's `R` code style guide.

# Optional: Acknowledgements
acknowledgements: |

# Optional: One or more keywords
keywords:

# Paper size for the document, values of letter and a4
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
# one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
# lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: pinp.bib

# Optional: Enable a 'Draft' watermark on the document
#watermark: true

# Customize footer, eg by referencing the vignette
footer_contents: "YourPackage Vignette"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{YourPackage}
  %\VignetteEngine{knitr::rmarkdown}
---

## Data Preparation 

There are two telematics data set from independent sensors: GPS in smartphone, OBDII. The two data sets are provided as two separate file. Here we assume that the two `json` files are extracted in the working directory, which means you have the following two files in the working directory:

> mobile_trips.json, obd2_trips.json

## Data load and structure

Since the two files are recorded as `json` format, `jsonlite` package will be used to load the data.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(jsonlite)
# read obd2 trip & mobile trip
obd2_data <- fromJSON("../obd2_trips.json")
mobile_data <- fromJSON("../mobile_trips.json")
```

```{r eval=FALSE}
library(jsonlite)

# read mobile trip & obd2 trip
mobile_data <- fromJSON("./mobile_trips.json")
obd2_data <- fromJSON("./obd2_trips.json")
```

Smartphone data set consisists of 44 trips and OBDII data set consists of 41 trips. The colum names of the each data set and the data types are as follows:

- OBDII data
  - trip_id (`char`), timestamp (`dbl`), speed (`int`)
- Mobile data
  - trip_id (`char`), created_at (`dbl`), timestamp (`dbl`), speed (`dbl`), accuracy (`int`)

## Visualization of sample trips

Fig \ref{fig:rawplot_obd2} shows the sample speed graph of one of the trips from OBDII data. As we can see, the trip was recorded for around 1,200 seconds and we do not know the unit for the recorded speed. By examing the data set littel bit, it can be easily realized that the first trip from each sources corresponds to each other as in Fig. \ref{fig:rawplot_obd2} and Fig \ref{fig:rawplot_mobile}.

```{r rawplot_obd2, fig.width=3, fig.height=3, cache=TRUE, echo=FALSE, fig.cap="A sample of speed graph of OBDII (the 1st trip)."}
library(ggplot2)
obd2_trip_sample1 <- obd2_data[[1]]
ggplot(obd2_trip_sample1, aes(timestamp, speed)) +
    geom_line() +
    labs(x = "Time (sec.)",
       y = "Speed (unknown)") +
    theme(legend.position="none")
```

```{r rawplot_mobile, fig.width=3, fig.height=3, cache=TRUE, echo=FALSE, fig.cap="A sample of speed graph of Smartphone (the 1st trip)."}
mobile_trip_sample1 <- mobile_data[[1]]
ggplot(mobile_trip_sample1, aes(timestamp  - timestamp[1], speed)) +
    geom_line() +
    labs(x = "Time (sec.)",
       y = "Speed (unknown)") +
    theme(legend.position="none")
```

This *pinp is not PNAS* template started when the introduction to
[Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) by \cite{PeerJ:Rcpp} 
was converted into this updated
[Rcpp Introduction](https://eddelbuettel.github.io/pinp/Rcpp-introduction.pdf)
vignette.  It is based on the
[pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template of the wonderful [rticles](https://cran.r-project.org/package=rticles) package by
\cite{CRAN:rticles}. The conversion from markdown to latex is facilitated by
[rmarkdown](https://cran.r-project.org/package=rmarkdown)
\citep{CRAN:rmarkdown} and [knitr](https://cran.r-project.org/package=knitr)
\citep{CRAN:knitr}. The underlying LaTeX macros are from
[pnas.org](http://www.pnas.org/site/authors/latex.xhtml).

The remainder of the document carries over from the corresponding [pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template document. but has been edited and updated to our use case.  A few specific tips follow. In general, for fine-tuning some knowledge of LaTeX is helpful.


## Author Affiliations 

Per common academic best practice, you can include your department,
institution, and complete address, with the ZIP/postal code, for each
author. Use lower case letters to match authors with institutions, as
shown in the example. Authors with an ORCID ID may supply this
information at submission.

## Document Options

We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermark to be added to each page;
- Line of custom text in subtitle (`date_subtitle`) suitable to give publication info of 
the draft, e.g. journal name in a post-print. 



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures. The `\begin{widetext}...\end{widetext}` environment as shown
in equation \ref{eqn:example} below is deprecated, but \LaTeX commands
`\onecolumn` and `\twocolumn` work fine.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

